---
title: "APPLIED ECOX"
output: html_document
---
```{r}
library(tseries)
library(forecast)
library(repr)
library(rugarch) # package for GARCH modelling
library(moments) # package for kurtosis estimation
library(aTSA) # package for ARCH-LM test, it also contains different version of adf.test

options(repr.plot.width=8, repr.plot.height=5) # set the plotting parameters
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(quantmod)
library(xts)
library(PerformanceAnalytics)
library(rugarch)
library('tseries')   #Time Series Analysis and Computational Finance 
library('forecast') #Forecasting Functions for Time Series and Linear Models 
library('repr') #Serializable Representations 

MDAXI <- as.xts(X_MDAXI)
MDAXI <- na.omit(MDAXI)
chartSeries(MDAXI)

```
```{r}
plot(MDAXI)
```

```{r}
#STATIONARITY
MDAXI <- as.xts(X_MDAXI)
MDAXI <- na.omit(MDAXI)
```


```{r}
par(mfrow = c(1,2))
acf(MDAXI, main = "ACF")
pacf(MDAXI, main = "PACF")
```
```{r}
tseries::adf.test(MDAXI)
```

```{r}
log_ret <- diff(log(MDAXI))
log_ret = na.omit(log_ret)

chartSeries(log_ret)
```
```{r}
plot(log_ret)
```

```{r}
tseries::adf.test(log_ret)
```
```{r}
par(mfrow = c(1,2))
acf(log_ret, main = "ACF")
pacf(log_ret, main = "PACF")
```
According to the output we see pattern for MA(1) process since autocorrelation drops towards 0 for second lag and higher while partial correlation shows only a slow decay. Let=s try to also use selection procedure based on AIC and BIC:
```{r}
# choose the best model based on the AIC or BIC
(fit_auto1 <- auto.arima(log_ret, ic = c("bic"), stationary = TRUE))
(fit_auto2 <- auto.arima(log_ret, ic = c("aic"), stationary = TRUE))
```
```{r}
# plot the dependence structure left in the residuals
resid <- fit_auto1$residuals # save the residuals

par(mfrow = c(1,2))
Acf(resid, main = "ACF")
Pacf(resid, main = "PACF")
```
```{r}
# use the Ljung-Box test to verify if the ACF at first 20 lags is jointly equal to zero
# parameter 'fitdf' to specify the number of degrees of freedom to be subtracted if x is a series of residuals (p+q+constant)
Box.test(fit_auto2$residuals,  lag = 20, type = c("Ljung-Box"), fitdf = length(fit_auto2$coef))
```


```{r}
# we can employ function 'checkresiduals' in order to run diagnostic for the residuals
checkresiduals(fit_auto2)
```
Looks like there is nothing left regarding the linear dependency in the residuals. They should be thus Gaussian white noise. Let's check the Gaussianity assumption. Let's look at the distribution of standardized residuals defined as and compare them with the standard normal distribution.
```{r}
resid_norm <- resid/sqrt(fit_auto2$sigma2)

par(mfrow = c(1,2))
hist(resid_norm, prob = TRUE, breaks = 100, xlab = "Normalized residuals", main="Histogram")
curve(dnorm(x, mean = 0, sd = 1), col="blue", lwd = 1, add = TRUE)
qqnorm(resid_norm)
qqline(resid_norm)
```


```{r}
# Let's compute the Jarque-Bera test for the normality of the residuals
jarque.bera.test(resid)
kurtosis(resid)
skewness(resid)
```
Test rejects the hypothesis that the residuals are normal - based on the estimated values of kurtosis and skuewness of the residuals. The rejection is mainly driven by the high kurtosis of the data - fat tails. There is relatively high chance of extreme events - important feature for risk management, portfolio selection etc.
```{r}
resid_squared<-resid^2
plot(resid_squared, ylab="Squared Residuals")
```
```{r}
par(mfrow=c(1,2))
Acf(resid_squared, lag.max = 12, main="ACF")
Pacf(resid_squared, lag.max = 12, main="PACF")
```
Dependence in the squared residuals suggests that there is some serial structure in the second moment of the series - which was also suggested by the volatility clusters observed in the plot. Solution: employ volatility models (GARCH type of models, etc.) to properly model the associated risk.
How to test whether there is some dependence in squared residuals?
Ljung-Box test on squared residuals:
```{r}
options(repr.plot.width=10, repr.plot.height=7)
# portmanteau-Q test (similar to Ljung-Box test) and ARCH-LM test:
(fit_manual <- arima(log_ret, order = c(0, 0, 1)))
arch.test(fit_manual)
```
We can see that the p-values of ARCH-LM test strongly reject the null hypothesis that there is no heteroskedasticity in the residuals. Let's perform the Ljung-Box test on squared residuals as well to confirm the result.
```{r}
Box.test(resid^2,  lag = 8, type = c("Ljung-Box"))
```
We can see that the ARCH effect is indeed present. So, in order to model it, we employ GARCH models to capture the time-varying nature of the volatility of the residuals. The general specifiaction of GARCH(p,q) iswhere the residual from the ARMA model is defined as
ð‘’ð‘¡=ðœŽð‘¡ðœ–ð‘¡
where  ðœ–ð‘¡  is iid standard normal random variable and  ðœŽð‘¡  is the time-varying volatility of the residual.
So, let's estimate simple ARMA(0,1)-GARCH(1,1) model on the log returns and print the estimation results.
```{r}
hist(residuals(fit_1n, standardize = TRUE),freq=FALSE, nclass=40, xlab="Standardized Residuals of ARMA(0,1)-GARCH(1,1)")
curve(dnorm(x, mean=0, sd=1), 
      col="darkgreen", lwd=2, add=TRUE, yaxt="n")
```

```{r}
# specify the model that we want to estimate:
# we can estimate the ARMA-GARCH model jointly using rugarch package
model_specification_1n <- ugarchspec(mean.model = list(armaOrder = c(0, 1)), 
                     variance.model = list(garchOrder = c(1, 1)),
                     distribution.model = "norm")

# fit the model
(fit_1n <- ugarchfit(spec = model_specification_1n, data = log_ret))

fit_1n

# in the results, robust standard errors are obtained using quasi-mle instead of MLE. These errors are more robust against violations of the distributional assumption

# alpha+beta<1 is the stability condition - stationary unconditional variance, if alpha1>1 the shock would have destabilizing effect causing the variance -> inf.
```
```{r}
model_specification_1s <- ugarchspec(mean.model = list(armaOrder = c(0, 1)), 
                     variance.model = list(garchOrder = c(1, 1)),
                     distribution.model = "std")

# fit the model
(fit_1s <- ugarchfit(spec = model_specification_1s, data = log_ret))

fit_1s
```
```{r}
model_specification_1g <- ugarchspec(mean.model = list(armaOrder = c(0, 1)), 
                     variance.model = list(garchOrder = c(1, 1)),
                     distribution.model = "ged")

# fit the model
(fit_1g <- ugarchfit(spec = model_specification_1g, data = log_ret))

fit_1g
```

```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_2n <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "eGARCH", garchOrder = c(1, 1)),
                     distribution.model = "norm") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_2n <- ugarchfit(spec = model_specification_2n, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_2n


```

```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_2s <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "eGARCH", garchOrder = c(1, 1)),
                     distribution.model = "std") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_2s <- ugarchfit(spec = model_specification_2s, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_2s

```
```{r}
hist(residuals(fit_2s, standardize = TRUE),freq=FALSE, nclass=40, xlab="Standardized Residuals of ARMA(0,1)-GARCH(1,1)")
curve(dnorm(x, mean=0, sd=1), 
      col="darkgreen", lwd=2, add=TRUE, yaxt="n")
```

```{r}
hist(residuals(fit_2n, standardize = TRUE),freq=FALSE, nclass=40, xlab="Standardized Residuals of ARMA(0,1)-GARCH(1,1)")
curve(dnorm(x, mean=0, sd=1), 
      col="darkgreen", lwd=2, add=TRUE, yaxt="n")
```

```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_2g <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "eGARCH", garchOrder = c(1, 1)),
                     distribution.model = "ged") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_2g  <- ugarchfit(spec = model_specification_2g, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_2g 
```


```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_3n <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)),
                     distribution.model = "norm") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_3n <- ugarchfit(spec = model_specification_3n, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_3n


```


```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_3s <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)),
                     distribution.model = "std") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_3s <- ugarchfit(spec = model_specification_3s, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_3s


```

```{r}
# you can specify some special GARCH models using ugarchspec function
model_specification_3g <- ugarchspec(mean.model = list(armaOrder = c(0, 1), include.mean = FALSE), 
                     variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)),
                     distribution.model = "ged") # 'norm' normal distribution, 'std' student-t distribution
# estimate the model
fit_3g <- ugarchfit(spec = model_specification_3g, data = log_ret)

# Explore the results
#coef(fit) # estimated parameters
fit_3g


```

Now, we can plot the estimated volatility 
```{r}
resid = as.numeric(residuals(fit_auto2, standardize = FALSE)) # getting residuals from the mean equation

options(repr.plot.width=10, repr.plot.height=4)
plot(abs(resid),type="l",col="black", ylab="Absolute residuals", xlab="Time") # absolute value of residuals from arma model
lines(array(sigma(fit_2s)),col="yellow") # estimated volatility, estimated sigma
```

```{r}
options(repr.plot.width=10, repr.plot.height=3)
# now we can check the standardized residuals from the model (residuals multiplied by sqrt(h_t)^-1) - scaling based on estimated volatility
# check ACF and PACF of the squared residuls: (are there remaining Garch effects?)
res <- residuals(fit3, standardize = TRUE)
sqr_res <- residuals(fit3, standardize = TRUE)^2

par(mfrow = c(1,2))
Acf(res, lag.max = 20, main="ACF standardized residuals")
Pacf(res, lag.max = 20, main="PACF standardized residuals")

par(mfrow = c(1,2))
Acf(sqr_res, lag.max = 20, main="ACF squared standardized residuals")
Pacf(sqr_res, lag.max = 20, main="PACF squared standardized residuals")

# this can be checked more formally using i.e. Ljung-Box test (see output of garch estimation)
```

```{r}
# check visually normality of the residuals
hist(residuals(fit3, standardize = TRUE),freq=FALSE, nclass=40)
curve(dnorm(x, mean=0, sd=1), 
      col="darkgreen", lwd=2, add=TRUE, yaxt="n")
# this can also be checked more formally (i.e. Jarque-Bera test)
```
Although, there are still some heavy tails present, the distribution is much more closer to the normal distribution than in the case of unconditional volatility! And this was achieved by the simple GARCH(1,1) model.

```{r}
#h0=normality of the data
# normality of residuals is still rejected, but the standardized residuals follow normal distribution more closely now.
jarque.bera.test(residuals(fit3, standardize = TRUE))
```



# Plot estimated variances
plot(variances, type = "l", col = c("red", "blue", "green"),
     main = "Estimated Variances for GARCH Models",
     xlab = "Date", ylab = "Variance")
legend("topright", legend = c("sGARCH-normal", "sGARCH-t", "eGARCH-t"), 
       col = c("red", "blue", "green"), lty = 1)
``
```{r}
colnames(log_ret) <- c('Returns')
colnames(X_MDAXI) <- c('Prices')
data <- merge.xts(X_MDAXI, log_ret)

```

```{r}
as.numeric(log_ret)[1]
```

```{r}
for (i in 2:nrow(data)){ # starting in day 2, 
  # since in day 1 is not possible; simple: the difference
  data[i, "vola"] <- as.numeric(log_ret)[i] - as.numeric(log_ret)[i-1]
}
```

```{r}
data$vola_sq <- NA 
for (i in 2:nrow(data)){ # starting in day 2, 
  # since in day 1 is not possible; simple: the difference
  data[i, "vola_sq"] <- data[i, "vola"]^2
}
```

```{r}
length(data$vola)
```
```{r}
sigma_est_1n <-as.vector(sigma(fit_1n)[-1])
```

```{r}
sigma_est_1s <-as.vector(sigma(fit_1s)[-1])
sigma_est_1g <-as.vector(sigma(fit_1g)[-1])
sigma_est_2n <-as.vector(sigma(fit_2n)[-1])
sigma_est_2s <-as.vector(sigma(fit_2s)[-1])
sigma_est_2g <-as.vector(sigma(fit_2g)[-1])
sigma_est_3n <-as.vector(sigma(fit_3n)[-1])
sigma_est_3s <-as.vector(sigma(fit_3s)[-1])
sigma_est_3g <-as.vector(sigma(fit_3g)[-1])
```

```{r}
sigma_sq_est_1n <-sigma_est_1n^2
sigma_sq_est_1s <-sigma_est_1s^2
sigma_sq_est_1g <-sigma_est_1g^2

sigma_sq_est_2n <-sigma_est_2n^2
sigma_sq_est_2s <-sigma_est_2s^2
sigma_sq_est_2g <-sigma_est_2g^2

sigma_sq_est_3n <-sigma_est_3n^2
sigma_sq_est_3s <-sigma_est_3s^2
sigma_sq_est_3g <-sigma_est_3g^2
```


```{r}
real_volatility <- as.vector(data$vola_sq[-1])
```


```{r}
real_volatility_sq <- real_volatility^2
```


```{r}
library(MLmetrics)
MSE1_1n <- MSE(sigma_sq_est_1n, real_volatility_sq)
MSE1_1s <- MSE(sigma_sq_est_1s, real_volatility_sq)
MSE1_1g <- MSE(sigma_sq_est_1g, real_volatility_sq)

MSE1_2n <- MSE(sigma_sq_est_2n, real_volatility_sq)
MSE1_2s <- MSE(sigma_sq_est_2s, real_volatility_sq)
MSE1_2g <- MSE(sigma_sq_est_2g, real_volatility_sq)

MSE1_3n <- MSE(sigma_sq_est_3n, real_volatility_sq)
MSE1_3s <- MSE(sigma_sq_est_3s, real_volatility_sq)
MSE1_3g <- MSE(sigma_sq_est_3g, real_volatility_sq)
```


```{r}
print(min(MSE1_1n,MSE1_1s,MSE1_1g,MSE1_2n,MSE1_2s,MSE1_2g,MSE1_3n,MSE1_3s,MSE1_3g))
```
MSE1 min je EGARCH normal distr.
```{r}
MSE1_2g 
MSE2_2g
MAE1_1s
MAE2_2g
```

```{r}
MSE2_1n <- MSE(sigma_est_1n, real_volatility)
MSE2_1s <- MSE(sigma_est_1s, real_volatility)
MSE2_1g <- MSE(sigma_est_1g, real_volatility)

MSE2_2n <- MSE(sigma_est_2n, real_volatility)
MSE2_2s <- MSE(sigma_est_2s, real_volatility)
MSE2_2g <- MSE(sigma_est_2g, real_volatility)

MSE2_3n <- MSE(sigma_est_3n, real_volatility)
MSE2_3s <- MSE(sigma_est_3s, real_volatility)
MSE2_3g <- MSE(sigma_est_3g, real_volatility)
```

```{r}
print(min(MSE2_1n,MSE2_1s,MSE2_1g,MSE2_2n,MSE2_2s,MSE2_2g,MSE2_3n,MSE2_3s,MSE2_3g))
```
MSE2 min je EGARCH normal distr.

```{r}
MAE1_1n <- mae(sigma_sq_est_1n, real_volatility_sq)
MAE1_1s <- mae(sigma_sq_est_1s, real_volatility_sq)
MAE1_1g <- mae(sigma_sq_est_1g, real_volatility_sq)

MAE1_2n <- mae(sigma_sq_est_2n, real_volatility_sq)
MAE1_2s <- mae(sigma_sq_est_2s, real_volatility_sq)
MAE1_2g <- mae(sigma_sq_est_2g, real_volatility_sq)

MAE1_3n <- mae(sigma_sq_est_3n, real_volatility_sq)
MAE1_3s <- mae(sigma_sq_est_3s, real_volatility_sq)
MAE1_3g <- mae(sigma_sq_est_3g, real_volatility_sq)
```

```{r}
print(min(MAE1_1n,MAE1_1s,MAE1_1g,MAE1_2n,MAE1_2s,MAE1_2g,MAE1_3n,MAE1_3s,MAE1_3g))
```
ALSO EGARCH NORMAL DISTR

```{r}
MAE2_1n <- mae(sigma_est_1n, real_volatility)
MAE2_1s <- mae(sigma_est_1s, real_volatility)
MAE2_1g <- mae(sigma_est_1g, real_volatility)

MAE2_2n <- mae(sigma_est_2n, real_volatility)
MAE2_2s <- mae(sigma_est_2s, real_volatility)
MAE2_2g <- mae(sigma_est_2g, real_volatility)

MAE2_3n <- mae(sigma_est_3n, real_volatility)
MAE2_3s <- mae(sigma_est_3s, real_volatility)
MAE2_3g <- mae(sigma_est_3g, real_volatility)
```

ALSO EGARCH NORMAL DISTR.
```{r}
print(min(MAE2_1n,MAE2_1s,MAE2_1g,MAE2_2n,MAE2_2s,MAE2_2g,MAE2_3n,MAE2_3s,MAE2_3g))
```